### 2013-10-30 analysis.Rmd
Begin editing for Moscow HSE LCSR seminar in November 2013

% Attributions for Poverty in Post-Socialist Countries
% Markus Kainu & Mikko Niemelä
% Jan 2013

<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>
<link href='http://fonts.googleapis.com/css?family=Alegreya+SC|Dosis:400,200' rel='stylesheet' type='text/css'>

******

<center>
<table>
<tr>
<td>[Etusivu](index.html)</td>
<td>[Analysis](analysis.html)</td>
<td>[Article](article2013.html)</td>
</tr>
</table>
</center>

*******

# Analysis