### 2013-10-30 code/multilevel_models.Rmd

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/code/")
library(knitr)
knit("multilevel_models.Rmd")
system("~/.cabal/bin/pandoc multilevel_models.md --number-sections --template=../tex_templates/dcolumn.latex -o multilevel_models.pdf")

system("~/.cabal/bin/pandoc -s multilevel_models.md -o multilevel_models.html")

#system("perl -pi -e 's/footnotesize/tiny/g' multilevel_models.tex")
```

<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>

# Final tables

```{rmultilevel_analysis, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, results='hide'}
library(lme4)
#########################################
#########################################
setwd("~/workspace/lits/attrib/attrib_year2013/code")

#########################################
load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")

summary(bisoc.1 <- lmer(pov.log.social ~ incsour3+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.2 <- lmer(pov.log.social ~ edu2+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.3 <- lmer(pov.log.social ~ income2+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.4 <- lmer(pov.log.social ~ past.diff+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.5 <- lmer(pov.log.social ~ future.diff+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.6 <- lmer(pov.log.social ~ crise+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.7 <- lmer(pov.log.social ~ group1+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.8 <- lmer(pov.log.social ~ uw.gini+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.9 <- lmer(pov.log.social ~ gdpchange07.10+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.10 <- lmer(pov.log.social ~ wbgi.vae+(1|cntry),
                         family = binomial, data=df))

summary(biind.1 <- lmer(pov.log.individual ~ incsour3+(1|cntry),
                        family = binomial, data=df))
summary(biind.2 <- lmer(pov.log.individual ~ edu2+(1|cntry),
                        family = binomial, data=df))
summary(biind.3 <- lmer(pov.log.individual ~ income2+(1|cntry),
                        family = binomial, data=df))
summary(biind.4 <- lmer(pov.log.individual ~ past.diff+(1|cntry),
                        family = binomial, data=df))
summary(biind.5 <- lmer(pov.log.individual ~ future.diff+(1|cntry),
                        family = binomial, data=df))
summary(biind.6 <- lmer(pov.log.individual ~ crise+(1|cntry),
                        family = binomial, data=df))
summary(biind.7 <- lmer(pov.log.individual ~ group1+(1|cntry),
                        family = binomial, data=df))
summary(biind.8 <- lmer(pov.log.individual ~ uw.gini+(1|cntry),
                        family = binomial, data=df))
summary(biind.9 <- lmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
                        family = binomial, data=df))
summary(biind.10 <- lmer(pov.log.individual ~ wbgi.vae+(1|cntry),
                         family = binomial, data=df))

summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
                      family = binomial, data=df))
summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+(1|cntry),
                      family = binomial, data=df))
summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+(1|cntry),family = binomial, data=df))
summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))

summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
                      family = binomial, data=df))
summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+(1|cntry),
                      family = binomial, data=df))
summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+(1|cntry),family = binomial, data=df))
summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
```

\newpage

```{rmultilevel_social, results='asis', echo=FALSE, eval=FALSE, cache=TRUE}
library(stargazer)
stargazer("Empty model"=soc.0,
          "Model sb1"=soc.1,
          "Model sb2"=soc.2,
          "Model sb3"=soc.3,
          "Model sb4"=soc.4,
          title="Logistic multilevel random intercept model for social blame type of explanation (Using stargazer-package (missing country level random effects)).")
```

```{rmultilevel_individual2_texreg, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
htmlreg(list(ind.0,
          ind.1,
          ind.2,
          ind.3,
          ind.4),
       use.packages=FALSE,
       include.pvalues=TRUE,
       caption="Logistic multilevel random intercept model for individual blame type of explanation (Using texreg-package (p-value stars missing)).", 
       caption.above=TRUE)
```

```{rmultilevel_individual, results='asis', echo=FALSE, eval=FALSE, cache=TRUE}
stargazer("Empty model"=ind.0,
          "Model ib1"=ind.1,
          "Model ib2"=ind.2,
          "Model ib3"=ind.3,
          "Model ib4"=ind.4,
          title="Logistic multilevel random intercept model for individual blame type of explanation. (Using stargazer-package (missing country level random effects))")
```

```{rmultilevel_social_texreg, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
htmlreg(list(soc.0,
       soc.1,
       soc.2,
       soc.3,
       soc.4),
       use.package=FALSE,
       include.pvalues=TRUE,
          caption="Logistic multilevel random intercept model for social blame type of explanation( Using texreg-package (p-value stars missing)).", caption.above=TRUE)
```

```{rbivariate_social, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(texreg)
htmlreg(list(bisoc.1,bisoc.2,
          bisoc.3,bisoc.4,
          bisoc.5,bisoc.6,
          bisoc.7,bisoc.8,
          bisoc.9,bisoc.10),
       use.packages=FALSE,
       sideways=TRUE,
       scriptsize=TRUE,
       include.pvalues=TRUE,
       caption="bivatiate social blame", caption.above=TRUE)
```

```{rbivariate_individual, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(texreg)
htmlreg(list(biind.1,biind.2,
          biind.3,biind.4,
          biind.5,biind.6,
          biind.7,biind.8,
          biind.9,biind.10),
       use.packages=FALSE,
       sideways=TRUE,
       scriptsize=TRUE,
       caption="bivatiate individual blame", caption.above=TRUE)
```