### 2013-10-31 loadClean.Rmd
Changing the data munging into vector algebra

% Loading and cleaning the data

- [back to index](index.html)

`generated at:`
```{r, echo=FALSE}
Sys.time()
```
---

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/")
library(kaRski)
knitpandoc("loadClean")
```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
```

# Loading the data from EBRD

```{rLoadClean1}
# FROM WEB
#temp <- tempfile()
#download.file("http://www.ebrd.com/downloads/research/surveys/lits2.dta", temp)
#library(foreign)
#lits2 <- read.dta(temp)

# FROM LOCAL DISK
load("~/workspace/data/lits/lits2.RData")
```

# Munging the data

## Country names & groupings

```{rLoadClean2}
# different country groupings
## remove whitespaces from country
library(stringr)
lits2$cntry <- str_trim(lits2$country_, side="right")
lits2$cntry <- factor(lits2$cntry)

lits2$socialism[lits2$cntry %in% c('France','Germany','Great Britain',
                                   'Italy','Sweden')]                   <- 'Market economy'

lits2$socialism[lits2$cntry %in% c('Czech Republic','Estonia','Hungary',
                                   'Bulgaria','Latvia','Lithuania',
                                   'Poland','Slovakia','Slovenia',
                                   'Romania','Albania','Bosnia',
                                   'Bulgaria','Croatia','Macedonia',
                                   'Montenegro','Kosovo','Serbia',
                                   'Armenia','Azerbaijan','Belarus',
                                   'Georgia','Kazakhstan','Kyrgyzstan',
                                   'Moldova','Mongolia','Russia',
                                   'Tajikistan','Ukraine','Uzbekistan')] <- 'Post-socialist'

summary(factor(lits2$socialism))

lits2$group_analysis[lits2$cntry %in% c('France','Germany',
                                        'Great Britain',
                                        'Italy','Sweden')] <- 'Western Europe'

lits2$group_analysis[lits2$cntry %in% c('Czech Republic','Estonia',
                                        'Hungary','Bulgaria',
                                        'Latvia','Lithuania',
                                        'Poland','Slovakia',
                                        'Slovenia','Romania')] <- 'CEE'

lits2$group_analysis[lits2$cntry %in% c('Armenia','Azerbaijan',
                                        'Belarus','Georgia',
                                        'Kazakhstan','Kyrgyzstan',
                                        'Moldova','Tajikistan',
                                        'Ukraine','Uzbekistan','Russia')] <- 'CIS'

summary(factor(lits2$group_analysis))

```

## Recoding the dependent variable

```{rLoadClean3}
summary(lits2$q309)
# reason for poverty
library(stringr)
lits2$q309 <- as.factor(str_replace(lits2$q309, 
                                    "Don't know", 
                                    "Dont know"))
# All options
lits2$q309 <- as.character(lits2$q309)
lits2$q309[lits2$q309 %in% 'Because of injustice in our society'] <- 'Social Blame'
lits2$q309[lits2$q309 %in% 'Because of laziness and lack of willpower'] <- 'Individual Blame'
lits2$q309[lits2$q309 %in% 'Because they have been unlucky'] <- 'Individual Fate'
lits2$q309[lits2$q309 %in% 'It is an inevitable part of modern life'] <- 'Social Fate'
lits2$q309[lits2$q309 %in% 'Not stated'] <- 'Not stated'
lits2$q309[lits2$q309 %in% 'Dont know'] <- 'Dont know'

lits2$q309 <- factor(lits2$q309, levels=c("Social Blame",
                                          "Individual Blame",
                                          "Individual Fate",
                                          "Social Fate",
                                          "Not stated",
                                          "Dont know"))

summary(lits2$q309)
# not stated and dont know coded NA

lits2$poverty <- lits2$q309
lits2$poverty[lits2$poverty %in% c('Not stated','Dont know')] <- NA
lits2$poverty <- factor(lits2$poverty)
summary(lits2$poverty)

## For logistic regressions
# $pov.log.social

lits2$pov.log.social[lits2$q309 == 'Social Blame'] <- 1
lits2$pov.log.social[lits2$q309 != 'Social Blame'] <- 0

summary(factor(lits2$pov.log.social))

# $pov_log_individual
###

lits2$pov.log.individual[lits2$q309 == 'Individual Blame'] <- 1
lits2$pov.log.individual[lits2$q309 != 'Individual Blame'] <- 0

summary(factor(lits2$pov.log.individual))

```

## Independent variables

```{rLoadClean4}
# $sex
lits2$sex <- lits2$q102_1
# $income

lits2$income[lits2$q227 == 1:3]  <- 'Low income'
lits2$income[lits2$q227 == 4:7]  <- 'Middle income'
lits2$income[lits2$q227 == 8:10] <-'High income'

lits2$income <- factor(lits2$income, 
                       levels=c("Low income",
                                "Middle income",
                                "High income"))

lits2$income2 <- factor(recode(lits2$q227, 
                              "1:5='Low'; 
                              6:10='High'; 
                              else=NA"))

lits2$income <- factor(lits2$income2, levels=c("Low","High"))

# median(lits2)
#svyquantile(~q227, d.df, quantile = c(0.25, 0.5, 0.75), ci = TRUE)
library(car)
lits2$q227_rec <- recode(lits2$q227, "-111:0=NA")
###

# $future
###

lits2$future <- factor(recode(lits2$q229, 
                              "1:3='a) low income'; 
                          4:7='b) middle income'; 
                              8:10='c) high income'; 
                              else=NA"))

lits2$q229_rec <- recode(lits2$q229, "-100:0=NA")

lits2$future.diff[lits2$q227_rec > lits2$q229_rec] <- 'worse'
lits2$future.diff[lits2$q227_rec <= lits2$q229_rec] <- 'same or better'
lits2$future.diff <- factor(lits2$future.diff)

###

# $past
###

lits2$past <- factor(recode(lits2$q228, 
                              "1:3='a) low income'; 
                          4:7='b) middle income'; 
                              8:10='c) high income'; 
                              else=NA"))

lits2$q228_rec <- recode(lits2$q228, "-111:0=NA")

lits2$past.diff[lits2$q227_rec < lits2$q228_rec] <- 'worse'
lits2$past.diff[lits2$q227_rec >= lits2$q228_rec] <- 'same or better'
lits2$past.diff <- factor(lits2$past.diff)

###

# $age
###

lits2$ageclass <- factor(recode(lits2$q104a_1, 
                              "1:29='29 or younger'; 
                          30:44='30 to 44'; 
                              45:59='45 to 59';
                              60:120='60 and older';
                                else=NA"))

lits2$ageclass <- factor(lits2$ageclass, levels=c("29 or younger","30 to 44","45 to 59","60 and older"))

lits2$age <- lits2$q104a_1

###

# $incsour tulonlähde
###

lits2$q226m_txt <- recode(lits2$q226m, 
                          "1='Salary or wages in cash or in kind';
                                2='Income from self-employment';
                          3='Sales or bartering of farm products';
                          4='Pensions';5='Benefits from the state';
                          6='Help from relatives or friends';7='Other'")

lits2$incsour3 <- factor(recode(lits2$q226m_txt, 
                                "c('Benefits from the state',
                                'Help from relatives or friends')='Dependent'
                                ;else='Not.dependent'"))

###

# $edu
###

lits2$edu2 <- factor(recode(lits2$q515, "c(1,2)='no or compulsory';
                           else='higher'"))

#################
# $crise - kriisikokemus
#################
lits2$q801 <- as.factor(str_replace(lits2$q801, "Don't know", "DONT KNOW"))
summary(lits2$q801)
lits2$q801 <- factor(lits2$q801, levels=c("A GREAT DEAL","A FAIR AMOUNT","JUST A LITTLE","NOT AT ALL","DONT KNOW","Refused"))

library(car)
lits2$crise <- recode(lits2$q801, "c('A GREAT DEAL','A FAIR AMOUNT')='Great or fair amount';
                       c('NOT AT ALL','JUST A LITTLE')='Little or not at all';
                       else=NA")

lits2$crise <- factor(lits2$crise, levels=c("Little or not at all",
                                            "Great or fair amount"))
summary(lits2$crise)
###

```

# Subsetting data for analysis

```{rLoadClean5, eval=TRUE}

###

## Data for the analysis
df <- subset(lits2, select=c("SerialID","weight","cntry","country0","q309",
                             "q309_rec","sex",
                             "ageclass","age",
                             "future.diff","past.diff",
                             "pov.log.social",
                             "pov.log.individual",
                             "socialism","incsour3",
                             "edu2",
                             "group_analysis",
                             #"group_general","group_fine",
                             "income2",
                             "crise",
                             "q227", # income
                             "q228", # menneisyys
                             "q229", # tulevaisuus
                             "q226m_txt", # tulonlähde
                             "q515", # edu
                             "q801" # crise,
                             ))
# filter out not wanted countries
df <- df[!is.na(df$group_analysis),] # valitaan vaan group1 -muuttujan maat
df$cntry <- factor(df$cntry)
# print data in csv-format

summary(df)

```

# Macro level data

## Summarising micro data at the country level

```{rLoadClean6, eval=TRUE}
###

library(reshape2)
library(survey)
d.df <- svydesign(id = ~SerialID, 
                  weights = ~weight, 
                  data = df)

summary309 <- data.frame(prop.table(svytable(~q309 + cntry, d.df), 2) * 100)
summary309 <- summary309[!is.na(summary309$Freq), ]

summary309w <- dcast(summary309, cntry ~ q309, value.var="Freq")
names(summary309w) <- c("cntry","social.blame","individual.blame","individual.fate","social.fate","not.stated","dont.know")

summary309_rec <- data.frame(prop.table(svytable(~q309_rec + cntry, d.df), 2) * 100)
summary309_rec <- summary309_rec[!is.na(summary309_rec$Freq), ]

summary309_recw <- dcast(summary309_rec, cntry ~ q309_rec, value.var="Freq")
names(summary309_recw) <- c("cntry","social.blame","individual.blame","individual.fate","social.fate")

```

## Data from Quality of Government institute

```{rLoadClean7, eval=TRUE}
#######
# Quality of governance
library(rustfare)
datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
                   'Kazakhstan','Tajikistan',
                   'Uzbekistan','Estonia','Latvia','Lithuania',
                   'Poland','Czech Republic','Slovakia','Hungary',
                   'Slovenia','Romania','Bulgaria',
                   'Italy','Germany','United Kingdom',
                   'Sweden','France'), 
       indicator = c("ffp_fsi", # Failed States Index
                     "wbgi_vae", # World Bank - Voice and Accountability
                     "undp_hdi", # UNDP Human Development Index
                     "fh_polity2", # Democracy (Freedom House/Polity)
                     "uw_gini"))
# subset the latest value per cntry per indicator
library(plyr)
datQogSum <- ddply(datQog, c("cname","indicator"), subset, year == max(year))

library(ggplot2)
ggplot(datQogSum, aes(x=year,y=value,color=cname)) + 
  geom_point() +
  geom_text(data = datQogSum,
            aes(x=year,y=value,color=cname,label=cname)) +
  facet_wrap(~indicator, scales="free") +
  theme(legend.position="none")

datQogSum.w <- dcast(datQogSum, cname ~ indicator, value.var="value")
datQogSum.w$cname <- recode(datQogSum.w$cname, "'United Kingdom'='Great Britain'")

datQogSum.w$group_fine <- recode(datQogSum.w$cname, "c('Armenia','Azerbaijan','Georgia') = 'Caucasus';
                         c('Russia','Ukraine','Moldova','Belarus')= 'CIS North';
                         c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')='Central Asia';
                         c('Estonia','Latvia','Lithuania')='Baltic States';
                         c('Poland','Czech Republic','Slovakia','Hungary','Slovenia',
                         'Romania','Bulgaria')='CEE EU';
                         c('Italy','Germany','Great Britain','Sweden','France')='West Europe';
                         else='other'")
datQogSum.w$group_general <- recode(datQogSum.w$cname,
                       "c('France','Germany','Great Britain',
                             'Italy','Sweden')='Western Europe';
                       c('Czech Republic','Estonia','Hungary','Bulgaria',
                       'Latvia','Lithuania','Poland',
                       'Slovakia','Slovenia','Romania')='CEE';
                       c('Armenia','Azerbaijan','Belarus','Georgia','Kazakhstan',
                       'Kyrgyzstan','Moldova','Tajikistan',
                       'Ukraine','Uzbekistan','Russia')='CIS';else=NA")

###

df.qog <- datQogSum.w
# underschores into dots for latex...
#names <- names(df.qog)
#library(stringr)
#names <- str_replace_all(names, "_",".")
#names(df.qog) <- names
library(gdata)
df.qog <- rename.vars(df.qog, from="cname", to="cntry")

```

## Data from World Bank

```{rLoadClean8, eval=TRUE}
## GDP change from WB
### We need iso2c countrycode
library(countrycode)
cntry <- levels(df$cntry) # countrynames
iso2c <- countrycode(cntry, "country.name", "iso2c")
iso2c # copy paste into WDI search
library(WDI)
# WDIsearch(string='gdp', field='name', cache=NULL)
wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), country = c("AL","AM","AZ","BY","BA","BG","HR",
                                                       "CZ","EE","FR","GE","DE","GB","HU",
                                                       "IT","KZ","KG","LV","LT","MK","MD",
                                                       "MN","ME","PL","RO","RU","RS","SK",
                                                       "SI","SE","TJ","TR","UA","UZ"), 
           start = 2007, end = 2010)

library(reshape2)
names(wb) <- c("iso2c","cntry","var","year")
wb.w <- dcast(wb, cntry ~ year, value.var="var")

names(wb.w) <- c("cntry","x2007","x2008","x2009","x2010")
wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
gdp <- wb.w[,c(1,6)]

gdp$cntry[gdp$cntry == 'Russian Federation'] <- 'Russia'
gdp$cntry[gdp$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
gdp$cntry[gdp$cntry == 'United Kingdom'] <- 'Great Britain'
gdp$cntry[gdp$cntry == 'Slovak Republic'] <- 'Slovakia'

```

## Merging all the data and saving for the analysis

```{rLoadClean9, eval=FALSE}

# yhdistetään nää aikaisempien datojen kanssa
macro <- merge(summary309w,gdp,by="cntry", all.x=TRUE)
macro <- merge(macro,df.qog,by="cntry")

df <- merge(df,macro,by="cntry")

rm(lits2)
rm(df.qog)
rm(df.cross.postsoc)
rm(gdp)
rm(t3)
rm(summary309)
rm(summary309_rec)
rm(d.df)
rm(wb)
rm(wb.w)
rm(cntry)
rm(iso2c)

save.image("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
#load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
```