### 2013-11-01 finalTables.Rmd

% Final regression ouput tables for the paper

- [back to index](index.html)

`generated at:`
```{r, echo=FALSE}
Sys.time()
```
---

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/")
library(kaRski)
knit("finalTables.Rmd")
system("pandoc -s finalTables.md -o finalTables.html --toc --number-section")

```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
```

**In progress..**

See old model outputs [here](file:///home/aurelius/workspace/lits/attrib/attrib_year2013/code/multilevel_models.html)