### 2013-11-09 article2013.Rmd
hieman epätoivoinen commit enne mikon tekstin lisäämistä

% Attributions for Poverty in Post-Socialist Countries[^*]
% Markus Kainu[^**];Mikko Niemelä[^***]
% October 2013

[^*]:*Work in progress

[^**]:PhD Student. Aleksanteri institute, University of Helsinki & Department of Social Research, University of Turku <a href="mailto:markuskainu@gmail.com">markuskainu@gmail.com</a>

[^***]:Senior Researcher. Social Insurance Institution of Finland, KELA a<a href="mailto:mikko.niemela@kela.fi">mikko.niemela@kela.fi</a>

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{roptsattribArticle, echo=FALSE}
opts_chunk$set(echo=FALSE,eval=TRUE,cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
```

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/")
library(knitr)
knit("article2013.Rmd")
system("~/.cabal/bin/pandoc article2013.md --number-sections -o article2013.pdf")
system("~/.cabal/bin/pandoc -s article2013.md --toc --number-sections -o article2013.html")
system("~/.cabal/bin/pandoc -f markdown -t docx article2013.md -o article2013.docx --standalone --toc --number-sections")
system("~/.cabal/bin/pandoc article2013.md -o article2013.epub --standalone --toc --number-sections")

# jos multilevelit mukana
#system("~/.cabal/bin/pandoc article2013.md --number-sections --template=tex_templates/dcolumn.latex -o article2013.tex")
#system("perl -pi -e 's/footnotesize/tiny/g' article2013.tex")
#system("pdflatex article2013.tex")

```

**Nootti**

*Paperi on vielä hieman vaiheessa ja suosittelen tutustumaan tutkimuksen analytiikkaan osoitteessa: [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/). Yritän parhaillaan saada monitasomallintamista paremmalle tolalle. Paperissa olevat mallit ovat alustavia ja pitävät sisällään sekä teoreettista että menetelmällistä epätarkkuutta.*

****

**Abstract**

*Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries. The study applies multilevel techniques in order to analyse the role of individual and country-level explanatory factors as determinants of poverty attributions. As country-level determinants, paper analyses the changes in country's economic performance during the period of global financial crises, income inequality and the state of democracy. As individual-level determinants, the paper tests to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty.*

\newpage

# Introduction

As Joseph Schumpeter (1942: 12) once argued, "attitudes are coins that do not readily melt". This means that normative expectations, justice beliefs or welfare perceptions are very hard to change. In this way, welfare attitudes can be associated with policy changes by highlighting public sentiments, which are normative background assumptions or ideas that constrain decision-making and institutional change by limiting the range of programme that decision-making elites are likely to perceive as acceptable and legitimate to both their constituents and themselves (Campbell, 2004: 96-100). On the other hand, "the worlds on welfare" literature on welfare attitudes has in turn emphasized that type of welfare policy affects public attitudes (Andress and Heien, 2001; Arts and Gelissen, 2001; Jaeger, 2006; Svallfors, 1997). In regard to public perceptions of the causes of poverty, in the marginal welfare-policy dominated liberal regime, the poor will be asked to fulfil much harder deservingness criteria than in a universal social democratic regime (van Oorschot 2006). In countries, in which welfare programmes are more marginal and less generous, explanations of poverty are more often based on individual morality, reflecting the strength of the dominant individualistic ideology, than they are in the Continental and Nordic welfare states – where external and structural explanations prevail (Albrekt Larsen, 2006: 47-56; Alesina and Gleaser, 2004).

The focus of this study is on the public perceptions of the causes of poverty in post-socialist countries. Since the early 1990s, the development of post-socialist welfare state has received abundant scholarly attention (e.g. Cerami and Vanhuysse, 2009, Inglot, 2008, Haggard and Kauffman, 2008; Cook, 2007; Aidukaite, 2004; Deacon 1992). As the decades under socialist regime have had an undeniable influence on societies, these studies have emphasized the role of institutional differences within region as well as pre-socialist historical origins of the welfare states. Now after 20 years of transition, as there has not been a visible convergence towards a common "post-socialist model", the attention has shifted even more on various institutional and historical explanations (Szikra and Tomka, 2009, Tomka, 2006, Cerami and Vanhuysse, 2009, Inglot, 2008).

The purpose of this article is to examine cross-country differences in attributions for poverty in Europe and Western Asia with special emphasis on post-socialist countries. Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasized clear differences between Western and Eastern European countries (Gallie and Paugam, 2002; Kallio and Niemelä, 2013; Lepianka, 2007; van Oorschot and Halman, 2000). East Europeans are more inclined to blame either the individual behavior or the flaws of the economic system of country than their Western counterparts. Hence, it seems that unlike in Western Europe traditionally opposite attitudinal dimensions, i.e. structural and individualistic explanations of poverty coincide with each other in East Europeans perceptions. However, there is shortage of comparative studies on attributions for poverty in particular focusing on the East Central European countries or on the whole post-socialist space. There is therefore a need for a more detailed and up-to-date analysis of the public perceptions of the causes of poverty in post-socialist countries.

The novelty of this study is that it covers wide range of countries from Europe and Western Asia. Thus, compared to previous studies, it provides more valid evidence on the differences in perceptions between the East and the West and within post-socialist regime as well as on the importance of the post-socialist legacy. In addition, the study applies multilevel techniques in order to analyze the role of individual and contextual level explanatory factors as determinants of poverty attributions. The article explores to what extent specific risks related to household financial circumstances at the individual level and the state of democracy, income inequality and changes in national economy at the contextual level are associated with different types of lay explanations of poverty. Finally, extensive country-selection of our data provides a novel opportunity to examine the effects of the financial crises both at the individual and contextual level and their association with attributions for poverty. At the individual level, the study focuses not only on the individuals’ current financial situation. Instead, the article provides an empirical example of how past experiences, current financial position as well as future expectations concerning individual’s economic situation are associated with perceptions of the causes of poverty.

The article is structured as follows: A brief overview of theoretical discourse on explanations of poverty and prior findings on cross-national differences in and contextual level determinants of attributions for poverty is presented in Section 2. Section 3 focuses on the self-interest approach as an individual level explanatory factor of the perceptions with particular interest of risk factors related to individual’s financial situation.  Research design, survey data and methods are described in detail in Section 4. The results of the study are presented in Section 5. Finally, the main conclusions, implications and limitations of the study are discussed in Section 6.

# Prior findings on cross-national differences

Pioneering empirical studies on lay explanations of poverty emerged at the early 1970s when Joe R. Feagin (1972; 1975) examined Americans' perceptions of the causes of poverty. He categorized reasons for poverty into three basic categories: individualistic, structural and fatalistic. Later, many factor analytic studies have given empirical support to Feagin’s categorization (e.g. Feather, 1974; Furnham, 1982; Hunt, 1996; Morçöl, 1997; Nasser et al., 2002; Niemelä, 2008). Yet some studies have expanded our understanding by incorporating more contemporary beliefs into the attributional scales. These works emphasize especially the relevance of cultural attributions for poverty (Bullock et al., 2003; Cozzarelli et al., 2001).

van Oorschot and Halman (2000) have suggested a four-tier typology (Table 1) in which individual blame emphasizes internal factors, such as undesirable behaviour of the poor, moral looseness and laziness. Individual fate type of explanation emphasizes that poverty can be caused by uncontrollable and inescapable internal factors. These are factors such as personal misfortune, illness or just bad luck. Social fate points out that poverty is caused by societal factors, but these factors are due to impersonal and unavoidable processes such as economic recessions. Finally, social blame sees poverty as the result of processes induced and controlled by the actions of certain groups and parties in society. Therefore, these groups and parties can be blamed for poverty.

![Dimensions and types of explanations of poverty](figure/nelikentta.png)

Prior empirical findings on attributions for poverty in Europe have shown that there are some differences between the West and the East of Europe. Western Europeans are more inclined to support social explanations for poverty than their Eastern counterparts. On the other hand, individual blame type of explanation is generally more popular in the East than in the West of Europe (Kallio and Niemelä, 2013; Kreidl, 2000; Lepianka, 2007; Lepianka et al., 2010; van Oorschot and Halman, 2000). Earlier results suggest, however, that there are variations both within and across groups of countries. Interestingly enough, the variation is highest within the East Central European countries (Kallio and Niemelä, 2013). Also Lepianka et al. (2010) come to the conclusion that further investigation is required due to rather large variation across countries. They showed that individual fate type of explanation is supported relatively less frequently in the East of Europe than in Western European countries. Regarding the variation within the Eastern European countries, they found that in some countries individual blame is strongly endorsed: especially the citizens of Czech Republic attribute poverty in internal terms (see also Lepianka, 2007). On the other hand, there are countries like Lithuania, Poland, Romania and Croatia in which social blame type of explanation is strongly supported. Finally, social fate type of explanation – living in poverty is explained as simply being part of modern progress – is endorsed especially in Russia.

National case studies from Estonia and Russia have given support to findings of cross-national studies. There is strong support for the idea that the poor have only themselves to blame for their poverty. In addition, both of these countries have strong support for social blame as well. Estonians are as likely to blame the flaws of the economic system as the individualistic reasons. In Russia, on the other hand, the economic system is blamed even more than the individuals themselves (Gorshkov and Tikhonova, 2006; Stephenson, 2000). These results are consistent with Lepianka's (2007: 127–135) comparative analysis of Hungary, Russia, Czech Republic and Estonia. Interesting result is that social blame coincides strongly with individual blame. This illustrates the split-consciousness of public perceptions, which means the coexistence of both dominant and potentially challenging beliefs, where the former represents dominant stratification ideology or value structure of a given country and the latter everyday stratification-related experience of an individual (Kluegel et al., 1995).

Based on qualitative poverty studies in Central Asia, Caucasus and Ukraine and Moldova Dudwick et al. (2003: 23) states that, generally, people blamed their own poverty on the failure of the Soviet state and the corruption, indifference, and incompetence of their new leaders. At the same time, however, they often attributed the poverty of others to individual failure, such as laziness, alcoholism, having too many children to support, or having too few children to provide for their old age. Emphazising the positive association between horizontal and vertical trust (see e.g. Rothstein, 2000), they also reported that many of the respondents had become very distrustful of government officials and many suspected that government officials of exploiting the political and economic turmoil for their own gain, at the expense of ordinary citizens. With extensive country selection, the data utilized in this study provides a marvellous possibility to examine the association between attributions for poverty and the context regarding political process, civil liberties and political rights. We can assume that in those countries where the state of democracy is low, citizens blame more likely society for poverty than in those countries where citizens have political freedom and power to influence socio-political issues.

There are also considerable differences in other aspects of socio-political context between examined countries. One of these aspects is the extent of income inequality which varies extensively between examined countries. Thus, this study examines the role of income inequality as a contextual-level explanatory factor of public perceptions of the causes of poverty. The link between inequality and attributions for poverty is not that easy to predict. One the one hand, the public in countries with greater inequality is more inclined to mistrust “others” (e.g. Uslander, 2002). Therefore, the uneven distribution of wealth should mean stronger support for the individualistic explanation for poverty. On the other hand, cross-national studies on attributions for poverty have shown that the support for the individual blame is lower in countries with lower level of income inequality (e.g. Gallie and Paugam, 2002; Kallio and Niemelä, 2013).

From the perspective of value structure, the role of "socialist or soviet legacy" in poverty attributions is worth of considering. Murthi and Tiongson (2008) reviewed large body of literature on East-West differences in preferences for redistribution of which vast majority supported the "socialist legacy". By socialist legacy they meant that "owing to the legacy of the past there is a strong preference for income equality and redistributive state spending in post-socialist countries". While testing the findings of previous studies on East-West differences, this study explores the effect of country groups with different socialist and post-socialist past. In order to explore the differences between market economies and post-socialist countries, but also within the "ex-soviet bloc", the study divides post-socialist countries into two groups: Central Eastern Europe (CEE) and Commonwealth of Independent States (CIS) countries. In order to examine the distinctiveness of post-socialist perceptions descriptive analyses include also a selection of Western European countries.

Besides the so called regime-effect, national-level economic conditions have been proved to be associated with welfare attitudes. This is particularly interesting in these times when the Western world is meeting the increasing economic turbulence. In regard to East-West differences as well as variation within post-socialist countries, countries examined in this study have faced global economic crises in distinct ways. While many Western and Eastern European countries have witnessed economic decline, the economic performance in Western Asia has been strong. Hence, by analyzing the role of changes in economic performance during the period of global financial crises, the research design of this article provides empirical evidence on the effects of crises to public perceptions of the causes of poverty. Prior studies have shown that increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune, 2007). In regard to explanations of poverty, Gallie and Paugam (2002:21-24) found that in the majority of European countries, there was a marked decline of support for the individual blame explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved. This result is confirmed by Kallio and Niemelä (2013). They found that the larger the economic growth in a country is, more likely people are to blame the poor themselves. Thus, we can assume that changes in country’s economic performance are associated with attributions for poverty regardless of country grouping.

# Risks factors as an individual-level determinants of perceptions

In regard to individual-level determinants, popular attitudes toward the welfare state have traditionally been assumed to be dependent on long-term class-related interest, short-term self-interest and the values and norms that have been internalized by the individuals in the society in question (Taylor-Gooby, 1985; Svallfors, 1995). More recent directions of comparative literature on welfare attitudes tend to also focus on the distribution of marketable capacities such as skill specificity (Iversen and Soskice, 2001) and resource/class-based risks (Cusack et al. 2006; Blomberg et al. 2012; Finseraas and Ringdal, 2012) as key to welfare attitudes. Thus, studies have assumed – and to some extent proved – that groups that have a greater risk of facing social problems or being dependent on welfare state benefits might perceive the government’s role in welfare issues in a distinct way compared to those who are not exposed to these risks.

Inspired by the studies which emphasise risks as the key mechanisms, this article tests to what extent factors related to financial hardship are associated – beyond the welfare regime – with perceptions of the causes of poverty. Following this line of reasoning, we are interested in whether those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks. Thus, the research design in regard to determinants of poverty attributions enables us to study from a fairly unexplored perspective the different mechanisms through which individuals who have a greater risk to experience material destitution perceive the causes of poverty.

The previous studies on individual-level determinants of poverty perceptions have focused on individual’s socio-demographic characteristics (e.g. Bullock, 1999; Cozzarelli et al., 2001; Furnham, 1982; Morcöl, 1997; Niemelä, 2008). Results have, however, shown that socio-demographic variables explain very little and they do not reveal clear or consistent patterns. On the other hand, there are studies which support the self-interest hypothesis that those who perceive themselves to be poor are more likely to blame society and social structures for poverty than people well above the poverty line (Niemelä, 2008; Saunders, 2002: 155–156). Yet, there are also contradictory findings indicating that personal experience of poverty might lead to individualistic explanations. For instance, claimants of public welfare seem to hold negative views about other recipients (Bullock, 1999; 2004; Golding and Middleton, 1982: 178).

Studies which have explored the importance of individual’s economic situation on attributions for poverty have focused on individual’s cross-sectional situation of financial circumstances. However, as suggested in the literature on welfare attitudes, not only the risk patterns but also the way people perceive their future risks might have an effect on attitudes toward the welfare state (Andress and Heien, 2001; Boeri et al., 2001; Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012). In their European comparison Blomberg et al. (2012) found that perceived future risks functioned as an intermediating factor between "objective" risk and welfare state attitudes, i.e. those who perceive their economic future as endangered were more in favour of state responsibility than those who do not expect their future situation to be difficult. Consequently, following the self-interest hypothesis, we can assume that those who perceive their future economic situation to be worse than now endorse more likely social blame type of explanation of poverty than those who do not expect difficulties in their future economic situation.

Not only the current situation or the perceived future but also the past experiences might be associated with individual’s evaluations (e.g. Kumlin, 2007). In regard to attributions for poverty, the effects of past experiences are – according to our knowledge – totally unexplored area of research. In this article we will analyse past experiences by two ways. First, we are interested in whether those whose current economic situation is lower than four years ago perceive the causes of poverty distinct way compared to those whose economic situation has improved or remained in the same level. Second, we will focus on the effects of the global financial crisis. We assume that economic crisis have affected especially those who are the most vulnerable groups in the society. Therefore, we will analyse whether those who report that the economic crisis has affected their household at least a fair amount perceive the causes of poverty differently than those who report that it has had no or only a little effect to their household. According to self-interest hypothesis, we assume that those whose economic situation has declined and those whose household has affected by economic crisis are more inclined to blame society for poverty than those who has not experienced economic problems.

In addition to economic situation, other indicators which measure a greater than average risk of facing more permanent social problems, are "transfer dependency" and low education. Transfer dependency refers to a situation that the respondent’s main source of income relies on private or social transfers (Blomberg et al., 2012; Fridberg and Kangas, 2008). Because of their close relation with welfare system, it can be assumed that welfare recipients perceive more likely social blame than indivual blame type of explanation of poverty compared to the rest of the population.

Low education refers to the lack of capabilities of an individual. Prior findings do not reveal consistent patterns regarding the effects of education. On the one hand, some studies have found the so-called inverted U-relationship which means that people with a middle level of education support the individualistic explanation, whereas people with a low or a high level of education are more likely to endorse structural explanations emphazising social inequalities and the flaws of the economic system of a given country (Feagin, 1972; cf. Furnham, 1982). On the other hand, there are studies which suggest that people with a lower level of education are more likely to favour individualistic explanations. Moreover, when the educational level increases, the likelihood of perceiving causes of poverty in individualistic terms decreases (Niemelä, 2008). It is therefore difficult to form any solid hypotheses as to whether low education as a social risk is associated with perceptions of the causes of poverty in a similar vein than other risk factors discussed above.

# Research design

Earlier studies have shown that attributions for poverty differ in general between the West and the East of Europe but the variation between East Central European countries is relatively large. There is, however, a shortage of comparative studies which have focused particularly on differences within post-socialist countries. In addition, most of the prior research is based on attitude surveys from the 1990s and most of them have restricted country selection. Therefore, as previous studies have suggested (Kallio and Niemelä, 2013; Lepianka et al., 2010), there is a need for detailed and more up-to-date analysis which would focus on the attributions for poverty within the post-socialist space. Hence, the general purpose of the study is to analyze public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.

The study utilizes the data from the second wave of Life in Transition Survey[^***2] (LiTS2), conducted by European Bank of Development and Reconstruction (EBRD) and  World Bank (WB) in 2010. The analyses cover 22 post-socialist countries and 4 Western European countries. The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample (for more detailed description of sampling and data collection, see ERBD, 2011). The survey had a special focus on how people’s lives were affected by the global economic crises and its aftermath.

[^***2]:[ebrd.com/pages/research/publications/special/transitionII.shtml](http://www.ebrd.com/pages/research/publications/special/transitionII.shtml)

The respondents were asked the following question which measures public perceptions of the causes of poverty: **In your opinion, what is the main reason why there are some people in need in our country today?**

- Because they have been unlucky
- Because of laziness and lack of willpower
- Because of injustice in our society
- It is an inevitable part of modern life
- Don't know
- Not stated

Thus, the question is a standard forced-choice question which is used also in other comparative surveys such as in World Values Survey and in Eurobarometer. According to van Oorschot and Halman (2000) different statements represent a four-tier typology of poverty explanations as follows: While the unlucky refers to individual fate, the laziness and lack of willpower represents individual blame. In addition, injustice in a society indicates social blame, whereas the view that poverty is simply part of modern life refers to social fate (see Table 1). Based on the findings of previous studies, we assume that in general post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations. However, we hypothesize also that findings indicate relatively large variation between countries.

As country-level determinants of perceptions, we will analyze, whether this variation can be explained by different regimes within post-socialist countries and the post-socialist legacy. The study divides post-socialist countries into two groups: Central Eastern Europe (CEE) and Commonwealth of Independent States (CIS). CEE includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania; and CIS[^****] Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia. The survey data enables us also to compare post-socialist countries to a selection of Western European nations (France, Germany, Italy and Sweden). Thus, in order to examine the distinctiveness of post-socialist perception, Western European countries are included in the descriptive analyses. (See Figure 1.)

![Countries used in the analysis](figure/mapplot.png)

[^****]:Georgia withdrawn its membership from CIS in 2008

Besides the country groups we will also analyze the role of changes in country’s economic performance during the period of global financial crises, income inequality and the state of democracy. As described in previous sections, we assume that in those countries where the economic crises has meant declining economic conditions, public are more likely to endorse social explanations for poverty than in countries where national economy has continued to grow despite the global economic turbulence. Based on the theoretical discussion on civil liberties, political rights and institutional trust, we assume that the public support to social blame type of explanation is high in those countries where the state of democracy is low. Due to contradictory findings on the direction of the effect of income inequality, it is more difficult to provide any straightforward hypothesis on the importance of income inequality.

As an individual-level determinants for the perceptions article test to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty, and whether these factors are associated with perceptions similar way regardless of the  groups of countries studied. Hereby, the research design aims at contributing to the theoretical reasoning regarding the importance of self-interest for attributions for poverty which assumes that those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks. Analyses include six individual-level predictors, namely transfer dependency, low education, perceived low income level, income-level compared to the past and expected change in income in the future as well as perceived effect of the financial crises. Variables are recoded as binomial where value 1 is coded for risk and 0 for absence of risk. Both contextual and individual-level variables are described in the Table 2.

\newpage

Table: Description of independent variables in the analyses.

-----------------------------------------------------------------------------
Variable                    Description
------------------------    -------------------------------------------------
Individual-level

Transfer dependency         Social or private transfers as main source of income

Low education               No or only compulsory level education

Perceived low income level  Perceived income level below the country median

Income compared to past     Perceived change in income compared to situation in four years ago

Income compared to future   Expected change in income over the next four years

Effect of financial crises  Whether respondents perceives that their household has suffered the financial crisis great or fair amount

Country-level

Ghange in total GDP 2007 – 2010   Relative change from 2007 to 2010

Gini coefficient                  Measures the distribution of income or consumption expenditure among individuals or households

Voice and Accountability          Indicators measuring the political process, civil liberties and political rights. These indicators measure the extent to which citizens of a country are able to participate in the selection of governments. Indicators measuring the independence of the media is also included.
Description
-----------------------------------------------------------------------------

An additional dataset of selected macro indicators were added to the LiTS2-dataset consisting of four variables: 1) change in total GDP in 2007 - 2010, 2) gini coefficient of economic inequality, 3) 2011 democracy index and 4) country grouping. Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), gini coefficients from [UNU-WIDER World Income Inequality Database](http://www.wider.unu.edu/research/Database/en_GB/database/) and Democracy Index from the [Economist Intelligence Unit](http://www.eiu.com/Default.aspx). Gini coefficient for each country is the latest available varying from 2002 to 2006. (see Appendix A)

TÄMÄ YLLÄ MAINITTU TEKSTI VOIDAAN KUITATA SÄRMÄSTI SILLÄ ETTÄ LISÄTÄÄN TUOHON TABLE 2:EN OIKEALLE PUOLELLE YKSI SARAKE OTSIKOLLA "DATA SOURCE" SIIHEN SITTEN INDIVIDUAL-LEVEL MUUTTUJIEN KOHDALLE JOKA RIVILLE LIFE IN TRANSITION SURVEY JA MAKROMUUTTUJIEN KOHDALLE TYYLIIN WORLD BANK 2013 JA SITTEN LÄHTEISIIN LÄHDEVIITEMERKINNÄT NÄILLE ELI WORLD BANK (2013) JA ONLINEVIITTAUS LEHDEN SÄÄTÖJEN MUKAISESTI.

The empirical section starts by presenting the results of descriptive analyses. As we assume that individual level perceptions are dependent both on individual and country-level contextual factors, it is recommended to use multilevel regression analysis for this task (Snijders \& Bosker, 2011).  Multilevel analysis requires hierarchical dataset and allows the estimation of country-level effects on individual level outcomes. The analysis is performed using lmer-function from lme4 package in R [^*****]. All of the models employed are so called multilevel logistic random intercept models at the level of individuals and countries.

[^*****]:R-code for data transformation and analysis is available at [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/)

TÄTÄ YLLÄ MAINITTUA ALAVIITETTÄ EI VARMAANKAAN SAA LAITTAA SUBMITTED-VERSIOON. SE VOIDAAN HALUTTAESSA LISÄTÄ SITTEN KUN PAPERI ON HYVÄKSYTTY LEHTEEN.

Our multilevel modelling focuses first in bivariate connections between two dependent variables and both individual and country level independent variables(See Appendix B). At the second step we include all individual level variables to a single model (models sb1 and ib1). Then we add each of the four country level variables one at the time (models sb2/ib2 - sb5/ib5). With limited number of countries we are obliged to use only few contextual variables. Tables will include both a fixed effects and a random effects part. The fixed effects part will include odds ratios and statistical significances of independent variables, while the random effects part will indicate the country-level variance and standard deviation, deviance and number of cases and groups. Full model summaries currently available at [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/).

# Results

## Descriptive statistics

Table 3 shows the proportion of the total number of respondents in a country which chose a particular explanation of poverty. We assumed that citizens in post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations. As expected results show relatively large variation within country groups. For the most of the post-socialist countries, social and individual blame types of explanation, i.e. injustice in a society and laziness of the poor, are the most popular poverty attributions. The highest support for social blame is found in Ukraine, Latvia, Lithuania, Hungary, Russia, Slovenia, Armenia, Moldova and Azerbaijan. In these countries majority of population blames the injustice in a society for poverty. Totally, in 19 of 21 examined post-socialist countries injustice in a society receives the highest support.

\newpage

```{rattribArticleDescTbl1, results='asis'}
source("code/dv_tables.R")
library(xtable)
print.xtable(xtable(tbl5, 
                    caption="Weighted population shares of dependent variable by country",),
              type="html",
              comment=FALSE, 
              include.rownames=FALSE,
             caption.placement="top")

```

Also, in most post-socialist countries the individual blame is the second most popular explanation of poverty. Exceptions here are Slovenia, Estonia and Georgia where citizens endorse societal explanations more likely than individual explanations. In Czech Republic social explanations as well as individual blame explanation receive public support quite evenly. Finally, the clearest exception in post-socialist regime is Uzbekistan where poverty is attributed strongly with internal terms.

In regard to differences between the East and the West of Europe, results show that there are some differences. Yet there are also substantial variation within Western Europe, which correspond earlier findings on cross-national differences in Western Europe (Lepianka et al., 2010; Kallio and Niemelä, 2010). Clearest differences between the West and the East is the stronger support for social fate in Western Europe. In Germany and Sweden external reasons for poverty receive the majority of support. On the other hand, attributions for poverty in France and Italy seems to be quite similar with post-socialist countries - blame types of explanations are endorsed more likely than fate types of explanations. Finally, the figures of Great Britain are the example of selective welfare policy dominated liberal regime in which the individual blame type of explanations is strongly supported (see also Albrekt Larsen, 2006).

Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty. Overall, the results indicate that for the most of the post-socialist countries two most popular explanations are individual and social blame types of explanations.Yet, the differences between social fate and individual blame types of explanations are in some countries very small. In general, Western Europeans (except Britons) do not support individual blame type of explanation as much as citizens in the East. Instead they emphasise more social reasons for poverty. Results also clearly emphasise that the idea that poverty is caused by just individual bad luck is the least popular perception of the reason for poverty.

## Determinants of perceptions

The results above indicate that there are cross-national differences between countries and within country groups which means that there are also some other contextual and individual-level factors which should explain the cross-national differences. Therefore, we will also analyse the importance of country-level economic, social and political conditions and individual-level risk factors related to financial hardship as determinants for attributions for poverty. Because the descriptive results showed that the majority of citizens in the post-socialist countries endorsed social and individual blame types of explanations, the subsequent analyses will focus only on these two explanations.

![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.png)

Analyses regarding the determinants of poverty explanations starts by bivariate associations between the dependent variables and three contextual variables (Figures 2, 3 and 4). Correlation coefficients at country group1 level are shown on top of each regression line.Figure 2 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crises (2007-2010). In regard to post-socialist countries there is substantial variation between countries in economic change. The general trend is quite clear: the less country has suffered from the financial crises the less there is support to social blame type of explanation (-.35) and more to individual blame (.48). In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries. However, this is greatly due to economic decline in Great Britain. Hence, without Great Britain we can not see any significant association between economic change and poverty perceptions in Western Europe.

![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.png)

There is also a substantial variation in the level of income inequality among post-socialist countries. Figure 3 shows that the higher level of income inequality is associated with higher level of social blame. As for individual blame type of explanation the picture is more puzzled, showing only weak connection between larger inequality and lower support for individual blame explanation. Again, the Western European countries differ with post-socialist countries. Cross-national differences in income inequality are much smaller in Western Europe than among post-socialist countries. There is a weak negative association between the social blame and income  inequality and a strong positive association between the individual blame and income inequality.

![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.png)

>Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media. [info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.png)

In regard to the level of democracy we can see great differences between and within groups of countries. Overall association is weak (.19 for social blame and -.44 for individual blame), but in CIS countries the higher level of democracy is clearly indicating higher support for social blame and lower support for individual blame type of explanation. In CEE and in Western Europe the variation in the state of democracy is much smaller than in CIS countries. In these countries the support for blame  types of explanations seems to vanish when the level of democracy increases. Consequently, both social and individual fate types of  explanations gain higher support in more democratic societies.

## Multilevel analysis

*There we some issues in converts regression outputs so the tables are evolving and can be browsed at: [http://research.muuankarski.org/attributions/code/multilevel_models.html](http://research.muuankarski.org/attributions/code/multilevel_models.html)*

****

<!--

As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables. Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.

-->

Table 4 and 5 present figures for multilevel logistic analysis. Calculating R² or residual standard deviation makes no sense in the case of logistic regression or other binomial models. Instead, we use \emph{deviance} as our primary statistical summary of logistic model fit. Deviance can be treated as similar to residual standard deviation for generalized linear models. Basically, deviance is a measure of error, and lower deviance implicates better fit to data. When a random predictor is added to model, we expect deviance to decrease, on average, by 1. When predictor is more informative deviance should decrease more than 1. (Gelman and Hill, 2007). As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.

Between-country variances are on the first rows in random effects, meaning the variance in dependent variable that is explained by country differences. Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation. Inclusion of individual level predictors decreases the variance in both models, but only moderately.

<!-- As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables. Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level. -->

Table 4 and 5 present figures for multilevel logistic analysis. Calculating R² or residual standard deviation makes no sense in the case of logistic regression or other binomial models. Instead, we use \emph{deviance} as our primary statistical summary of logistic model fit. Deviance can be treated as similar to residual standard deviation for generalized linear models. Basically, deviance is a measure of error, and lower deviance implicates better fit to data. When a random predictor is added to model, we expect deviance to decrease, on average, by 1. When predictor is more informative deviance should decrease more than 1. (Gelman and Hill, 2007). As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.

Between-country variances are on the first rows in random effects, meaning the variance in dependent variable that is explained by country differences. Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation. Inclusion of individual level predictors decreases the variance in both models, but only moderately.

Looking at the deviance for the empty models both in tables 4 and 5, we can see that there is "less" to be explained for in the model for individual blame than for social blame. When adding the individual level predictors to the models, the deviance decreases by 8512 for social blame and 6636 for individual blame. In both model sb1 and ib1 we have four statistically significant predictors. Low education and worsened household income level do not contribute in explaining neither of the outcomes. Perceived low income level has the highest coefficient of our binary individual level predictors for explaining social blame type of explanation. Coefficient of .368 can be translated so that if respondent perceives belonging to low income group he or she is 9.2\% more likely to blame society for poverty than a non-low income individual. Expected worsening in household income and financial crises experince predict social blame almost in similar manner. All our individual level predictors but low education predict lower probability for supporting individual blame (model ib1). Being greatly affected by financial crisis makes one 11\% less likely to support for individual blame type of explanation.

When adding in the country grouping (CEE/CIS) we can observe decrease close to -1 in deviance in both models sb2 and id2. Country grouping in itself is not significant in our models, whereas gini coefficient has a weak connection to individual blame type of explanation, so that individual blame is supported more in more unequal countries, as we assumed based on studies about the association between generalised trust and equality (Uslander 2002). Inclusion of change in national GDP between 2007 - 2010 substracts the deviance by 14 for social blame and 9 for individual blame model. Change in GDP becomes the most significant predictor of our country level variables for both type of explanation. The better the national economy has performed during the financial crises, the more probably is the support for individual blame type of explanation. Outcome is very much align with the findings of Gallie and Paugam (2002) and Kallio and Niemelä (2010). Our fourth contextual variable, democracy index, is significant in explaining the individual blame in a way that higher level of democracy implies lower support for individual blame type of explanation. Coefficient plots from both full models summarizes the estimates for logistic regression coefficients at 50\% and 90\% intervals in figure 7.

As for explaining the between-country variation the inclusion of country grouping adds a very little. Gini coefficient reduces variation from .24 to .19 for social blame, but it is the inclusion of GDP change that has the greatest effect. Between-country variation  to .12 for social blame and .04 for individual blame. In the final models the inclusion of democracy index results country-level variances at .09 and .02 for social and individual type of explanation, respectively. In other words, only 9\% and 2\% of the variation in the final models are due to country differences at final stage.

# Conclusions

To sum up, the results indicate that in general social blame type of explanation is clearly the most popular explanation in post-socialist countries. Social blame is followed by individual blame, social fate and individual fate types of explanations. Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty in the European post-socialist countries. In regard to differences between the East and the West of Europe, results show that Western Europeans emphasise more social reasons for poverty.

As assumed results also emphasised large between-country differences among post-socialist countries. By employing a multilevel approach we wanted to shed light especially on this issue. To begin with, we found that differences between countries are responsible for almost 30 percent of the variation in the support for social blame and 12 percent for individual blame type of explanation. Our contextual level predictors decreased the variance down to 9 percent for social and 2 percent for individual type of explanation. Yet, findings showed that the membership in the CIS and CEE country group was not significant in this respect.

As we hypothesised changes in country’s economic performance during the period of global financial crises were associated with attributions for poverty. In fact, changes in economic performance was the most significant predictor of our country-level variables for both social and individual blame types of explanations: the better the economy had performed the more support gained the individual blame type of explanation. And on the other hand, increased financial strain is associated with stronger support for the view that poverty is caused injustices of society.

Also, the level of democracy had a substantial impact. Yet, the results were to some extent against our hypothesis. Contrary to our assumption, citizens in those countries where the state of democracy is low do not blame more likely society than citizens in countries where they have political freedom and power to influence socio-political issues. Instead, findings indicate that the lower level of democracy was associated with stronger support for individual blame type of explanation. Social blame type of explanation was, in turn, endorsed more likely in countries with higher level of democracy.

Results of the individual-level determinants of perceptions supported the self-interest hypothesis which states that those who have financial difficulties perceive the causes of poverty differently than those who are not exposed to poverty-related risks. Findings indicate that those whose main source of income rely mainly on social or private transfers, those who perceive their income level to be lower than country’s median, those who perceive their future economic situation to be worse than now as well as those who perceive that the global financial crises has affected negatively to their household’s economic situation endorse more likely external (social blame) than internal (individual blame) reasons for poverty. Thus, analyses showed that that it would be fruitful to include a perspective of risk groups when studying attitudes toward poverty, and also to include future expectations in the equation (also Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).

This study focused only on poverty-related risks as individual-level determinants. There are, of course, other variables which might be associated with attributions even more than risk factors. For instance, another line of research have emphasised the role of other beliefs, attitudes and values rather than interest (e.g. Cozzarelli et al., 2001). Treating the values, attitudes and other perceptions as determinants for attributions of poverty is, however, problematic because it is difficult to prove the causal relationship between attributions and other attitudes. This does not mean that they are irrelevant. Conversely, we argue that future research should try to investigate to what extent high levels of social and individual blame are associated with wider welfare attitudes in post-socialist countries. In order, for example, to explore policy implications more thoroughly, there is a need for future research to focus on the relationship between attributions for poverty and the legitimacy of specific types of welfare policies in post-socialist countries.

As a methodological contribution, findings showed interesting differences between the East and the West regarding the effects of contextual variables. Even though the selection of Western European countries in this study included countries from different welfare regimes, the differences between countries in economic change, democracy and income inequality are quite modest compared especially to CIS countries. Consequently our results showed more modest and in some cases opposite associations between dependent and contextual variables among Western European countries than among post-socialist group.

Thus, the finding hold critical implications for future research. On the one hand, in order to examine the effects of contextual factors, comparative social research should try to utilize comparative data encompassing more diverse cases than, for instance, conventional EU17 or OECD countries. On the other hand, the well-known limitation of this kind of variable-oriented approach is that larger number of cases may pose also more obstacles to making interpretive statements about specific cases or even about categories of cases (Ragin, 1987). For instance, in regard to policy implications of research on welfare attitudes, broadening the country selection from Western democracies (in some cases) to more authoritarian societies raises a question whether public attitudes have similar constraining role in policy-making that have been assumed and to some extent proved to be in the policy-making of Western democracies. This kind of research question calls also case-oriented qualitative research methods.

# References

Aidukaite, Jolanta (2004) The Emergence of the Post-Socialist Welfare State - The Case of the Baltic States: Estonia, Latvia and Lithuania, Södertörn University College, Sweden.

Albrekt Larsen, Christian (2006) The Institutional Logic of Welfare Attitudes. How Welfare Regimes Influence Public Support. Hampshire: Ashgate.

Alesina, Alberto and Glaeser, Edward L. (2004) Fighting Poverty in the US and Europe. A World of Difference. Oxford: Oxford University Press.

Andress, Hans-Jürgen and Heien, Thorsten (2001) Four Worlds of Welfare State Attutudes? A Comparison of Germany, Norway, and the United States. European Sociological Review 17 (4), 337–356.

Arts, Wil and Gelissen, John (2001) Welfare States, Solidarity and Justice Principle: Does the Type Really Matter. Acta Sociologica 44 (4), 283–299.

Blomberg, Helena and Kallio, Johanna and Kangas, Olli and Kroll, Christian and Niemelä, Mikko (2012) Attitudes among High-Risk Groups. In Svallfors, Stefan (ed.) Contested Welfare States. Welfare Attitudes in Europe and Beyond. Stanford University Press: Stanford, CA, 58–80.

Boeri, T., Börsch-Supan, A. and Tabellini, G. (2001) Would You Like to Shrink the Welfare State? A Survey of European Citizens. Economic Policy 16, 7–50.

Bullock, Heather E. (1999) Attributions for Poverty: A Comparison of Middle-Class and Welfare Recipient Attitudes. Journal of Applied Social Psychology 29 (10), 2059–2082.

Bullock, Heather E. (2004) From the Front Lines of Welfare Reform: An Analysis of Social Worker and Welfare Recipient Attitudes. The Journal of Social Psychology 144 (6), 571-588.

Bullock, Heather E. and Williams, Wendy R. and Limbert, Wendy M. (2003) Predicting Support for Welfare Policies: The Impact of Attributions and Beliefs about Inequality. Journal of Poverty 7 (3), 35–56.

Campbell, John L. (2004) Institutional Change and Globalization. Princeton and Oxford: Princeton University Press.

Cerami, Aifio and Vanhuysse, Pieter (2009) Post-Communist Welfare Pathways. Theorizing Social Policy Transformations in Central and Eastern Europe. Hampshire: Palgrave Macmillan.

Cook, L. J. (2007) Postcommunist Welfare States: Reform Politics in Russia and Eastern Europe, Cornell University Press.

Cozzarelli, Catherine and Wilkinson, Anna V. and Tagler, Michael J. (2001) Attitudes Toward the Poor and Attributions for Poverty. Journal of Social Issues 57 (2), 207-227.

Cusack, T., Iversen, T. and Rehm, P. (2006) Risks at Work: The Demand and Supply Sides of Government Redistribution. Oxford Review of Economic Policy 22, 365–389.

Dudwick, N., Gomart, E., and Marc, A. (2003) When Things Fall Apart: Qualitative Studies of Poverty in the Former Soviet Union. World Bank Publications.

Economicst Intelligence Unit (2001) Democracy index 2011. Democracy under stress. Downloaded at: \href{http://www.eiu.com/democracyindex2011}{www.eiu.com/democracyindex2011}

Feagin, Joe R. (1972) Poverty: We Still Believe That God Helps Those Who Help Themselves. Psychology Today 6 (2), 101–129.

Feagin, Joe R. (1975) Subordinating the Poor: Welfare and American Beliefs. NJ Prentice Hall: Upper Saddle River.

Feather, Norman T. (1974) Explanations of Poverty in Australian and American Samples: The Person, Society, or Both? Australian Journal of Psychology 26 (3), 199–216.

Fridberg, Torben and Kangas, Olli (2008) The Welfare State, Poverty and Social Exclusion. In Ervasti, H., Fridberg, T., Hjerm, M. and Ringdal, K. (eds.) Nordic Social Attitudes in a European Perspective. Edward Elgar: Cheltenhamn, 22–47.

Furnham, Adrian (1982) Why are the poor always with us? Explanations for poverty in Great Britain. British Journal of Social Psychology 21(3), 311–322.

Gallie, Duncan and Paugam, Serge (2002) Social Precarity and Social Integration. Report for the European Commission Based on Eurobarometer 56.1. Brussels: European Commission Directorate-General Employment.

Giczi, Johanna and Sik, Endre (2009) Trust and Social Capital in Europe. In TARKI European Social Report 2009. TARKI: Budapest, 63-81.

Golding, Peter and Middleton, Sue (1982) Images of Welfare. Press and Public Attitudes to Poverty. Oxford: Basil Blackwell.

Gorshkov, Mikhail K. and Tikhonova, Natalia E. (2006) Wealth and Poverty in the Perceptions of Russians. Sociological Research 45 (1), 27–40.

Hunt, Matthew O. (1996) The Individual, Society, or Both? A Comparison of Black, Latino, and White Beliefs about the Causes of Poverty. Social Forces 75 (1), 293–322.

Inglot, T. (2008). Welfare States in East Central Europe, 1919-2004 (1st ed.). Cambridge University Press.

Iversen, T. and Soskice, D. (2001) An Asset Theory of Social Policy Preferences. American Political Science Review 95, 875–893.

Jaeger, Mads Meier (2006) Welfare Regime and Attitudes Towards Redistribution: The Regime Hypothesis Revisited. European Sociological Review 22 (2), 157–170.

Kallio, Johanna and Niemelä, Mikko (2010) Public Perceptions of the Causes of Poverty: Multilevel Analysis of 26 European Nations. Unpublished manuscript.

Haggard, S. and Kaufman, R. R. (2008) Development, Democracy, and Welfare States: Latin America, East Asia, and Eastern Europe, Princeton Univ Pr.

Kluegel, James R. and Mason, David S. and Wegener, Bernd (eds) (1995) Social Justice and Political Change: Public Opinion in Capitalist and Post-communist States, New York: Aldine De Gruyter.

Kouvo, Antti and Kankainen, Tomi and Niemelä, Mikko (2012) Welfare Benefits and Generalized Trust in Finland and Europe. In Ervasti, Heikki and Goul Andersen Jörgen and Fridberg, Torben and Ringdal, Kristen (eds.) The Future of the Welfare State. Social Policy Attitudes and Social Capital in Europe. Edward Elgar: Cheltenham, 195-213.

Kreidl, Martin (2000) Perceptions of Poverty and Wealth in Western and Post-Communist Countries. Social Justice Research 13 (2), 151–176.

Kumlin, Staffan (2007) Overload or Undermined? European Welfare States in the Face of Performance Dissatisfaction. In Svallfors, Stefan (ed.) The Political Sociology of the Welfare State: Institutions, Social Cleavages, and Orientations. Stanford University Press: Stanford, CA, 80–116.

Lepianka, Dorota (2007) Are the Poor to be Blamed or Pitied? A Comparative Study of Popular Attributions in Europe. Tilburg: Tilburg University.

Lepianka, Dorota and Gelissen, John and van Oorschot, Wim (2010) Popular Explanations of Poverty in Europe. Effects of Contextual and Individual Characteristics across 28 European Countries. Acta Sociologica 53 (1), 53–72.

Morçöl, Göktuğ (1997) Lay Explanations for Poverty in Turkey and Their Determinants. Journal of Social Psychology 137 (6) 728–738.

Murthi, M., & Tiongson, E. R. (2008). Attitudes to Equality: The Socialist Legacy Revisited. SSRN eLibrary. Retrieved from http://papers.ssrn.com/sol3/papers.cfm?abstract_id=1098627

Nasser, Ramzi and Abouchedid, Kamal and Khashan, Hilal (2002) Perceptions of the Causes of Poverty Comparing Three National Groups: Lebanon, Portugal, and South Africa. Current Research in Social Psychology 8 (7), 101–119.

Niemelä, Mikko (2008) Perceptions of the Causes of Poverty in Finland. Acta Sociologica, 51 (1), 23–40.

Nilson, Lina B. (1981) Reconsidering Ideological Lines: Beliefs about Poverty in America. The Sociological Quarterly 22 (4), 531-548.

Ragin, Charles C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies. Berkeley: University of California Press.

Rothstein, Bo (2000) Trust, Social Dilemmas and Collective Memories. Journal of Theoretical Politics, 12 (4), 477–501.

Rothstein, Bo and Stolle, Dietlind (2008) The State and Social Capital: An Institutional Theory of Generalized Trust. Comparative Politics 40 (4), 441-459.

Saunders, Peter (2002) The Ends and Means of Welfare. Coping with Economic and Social Change in Australia. Cambridge: Cambridge University Press.

Schumpeter, Joseph A. (1942) Capitalism, Socialism & Democracy. New York: Harper & Brothers.

Szikra, D., & Tomka, B. (2009). Social Policy in East Central Europe: Major Trends in the Twentieht Century. In Post-communist Welfare Pathways (Vol. 2009).

Stephenson, Svetlana (2000) Public Beliefs in the Causes of Wealth and Poverty and Legitimization of Inequalities in Russia and Estonia. Social Justice Research 13 (2), 83–100.

Svallfors, Stefan (1997) Worlds of Welfare and Attitudes to Redistribution: A Comparison of Eight Western Nations. European Sociological Review 13 (3), 283–304.

Tomka, B. (2006). East Central Europe and the European Social Policy Model: A Long-Term View. East European Quarterly, 40(2), 135.

van Oorschot, Wim and Halman, Loek (2000) Blame or Fate, Individual or Social? An International Comparison of Popular Explanations of Poverty. European Societies 2 (1), 1–28.

Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterly 37 (3), 413–428.