### 2013-11-11 mapPlots.Rmd

% Spatial visualization

- [back to index](index.html)

`generated at:`
```{r, echo=FALSE}
Sys.time()
```
---

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/")
library(knitr)
knit("mapPlots.Rmd")
system("pandoc -s mapPlots.md -o mapPlots.html --toc --number-section")
```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=15,fig.height=11)
```

# Country grouping

```{rmapPlotsLoadData}
library(RColorBrewer)
library(mapproj)
library(maptools)
library(rgdal)
load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 
                                layer="lits_all_simple"))

map$NAME <- as.character(map$NAME)
map$NAME[map$NAME == "United Kingdom"] <- 'Great Britain'
map$group_analysis[map$NAME %in% c('France','Germany',
                                        'Great Britain',
                                        'Italy','Sweden')] <- 'Western Europe'

map$group_analysis[map$NAME %in% c('Czech Republic','Estonia',
                                        'Hungary','Bulgaria',
                                        'Latvia','Lithuania',
                                        'Poland','Slovakia',
                                        'Slovenia','Romania')] <- 'CEE'

map$group_analysis[map$NAME %in% c('Armenia','Azerbaijan',
                                        'Belarus','Georgia',
                                        'Kazakhstan','Kyrgyzstan',
                                        'Moldova','Tajikistan',
                                        'Ukraine','Uzbekistan','Russia')] <- 'CIS'
map$NAME <- as.factor(map$NAME)

# shapefile into data.frame
map@data$id <- rownames(map@data)
library(ggplot2)
map.points <- fortify(map, region="id")
map.df <- merge(map.points, map@data, by="id")
# only east of longitude -25
map.df <- subset(map.df, long >= -15 & long <= 80)
# names for countries in analysis (excluding others)
namedata <- map.df[!is.na(map.df$group_analysis),]
cnames <- stats:::aggregate.formula(cbind(long, lat) ~ NAME, data=namedata, mean)
cbPalette <- c("#999999", "#E69F00",
               "#56B4E9", "#009E73",
               "#F0E442", "#0072B2",
               "gray90", "#CC79A7")
```

```{rmapPlotsGroup1}
ggplot(map.df, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
  scale_fill_manual(values=cbPalette) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
  coord_map(project="orthographic") +
  theme(legend.position="top")
```

# Dependent variables

**The darker the higher the support**

## Support for Social blame

```{rmapPlotsDV1}
### Faceted showing population shares of explanations
library(scales)
plotMap <- merge(map.df,macro,by.x="NAME",by.y="cntry")
plotMap <- plotMap[order(plotMap$order),]

ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = socialBlame)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

## Individual blame

```{rmapPlotsDV2}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = individualBlame)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

## Social fate

```{rmapPlotsDV3}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = socialFate)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

## Individual fate

```{rmapPlotsDV4}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = individualFate)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

## Not stated

```{rmapPlotsDV5}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = not.stated)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

## Dont know

```{rmapPlotsDV6}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = dont.know)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

# Independent variables

## Individual level

## Macro level

**Blue is "positive", red is "negative"**

### GDP change

```{rmapPlotsIvMacro1}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = gdpchange07.10)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low=muted("red"), high=muted("blue")) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

### Gini coefficient

```{rmapPlotsIvMacro2}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = uw_gini)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low=muted("blue"), high=muted("red"), midpoint=33.55) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```

### Voice and Accountability

```{rmapPlotsIvMacro3}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = wbgi_vae)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low=muted("red"), high=muted("blue")) +
  theme(legend.position="top") +
  theme(legend.text = element_text(size=14))
```