### 2013-11-21 compareWaves.Rmd
Changed the toc font size in css

# Comparing waves 2006 and 2010

```{r, eval=FALSE, echo=FALSE}
library(knitr)
library(markdown)
knit2html('compareWaves.Rmd', 
          stylesheet="~/workspace/web/css/rmarkdown.css",
          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
```

```{rcompload}
load("~/workspace/lits/attributions/data/lits.RData")
```

# Dependent variable

## Original coding

### By country

```{rcompareDvCntryOrig, fig.height=16}
dfPlot <- macro[,1:8]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("cntry","year"))
library(ggplot2)
ggplot(dfPlotW, aes(x=cntry,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free_x") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))
```

### By country group

```{rcompareDvCgOrig, fig.height=16}
dfPlot <- macro[,c(2:8,16)]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("group_general","year"))
library(ggplot2)
ggplot(dfPlotW, aes(x=group_general,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free_x") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))
```

## Only the four values

### By country

```{rcompareDvCntryRec, fig.height=16}
dfPlot <- macro[,c(1:2,9:12)]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("cntry","year"))
library(ggplot2)
ggplot(dfPlotW, aes(x=cntry,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free_x") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))
```

### By country group

```{rcompareDvCgRec, fig.height=16}
dfPlot <- macro[,c(2,9:12,16)]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("group_general","year"))
library(ggplot2)
ggplot(dfPlotW, aes(x=group_general,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free_x") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))
```

# Independent variables

## Individual level

### Transfer dependency (share of dependent hh's)

```{rcompareIvTransfer}
library(survey)
d.df10 <- svydesign(id = ~SerialID, weights = ~weight, data = df)
d.df06 <- svydesign(id = ~SerialID, weights = ~weight, data = df2)

dat10 <- data.frame(prop.table(svytable(~incsour3 + cntry, d.df10), 2) * 100)
names(dat10) <- c("Var","cntry","Freq")
dat10$year <- 2010
dat06 <- data.frame(prop.table(svytable(~incsour3 + cntry, d.df06), 2) * 100)
names(dat06) <- c("Var","cntry","Freq")
dat06$year <- 2006
datPlot <- rbind(dat06,dat10)

datPlot2 <- datPlot[datPlot$Var == "dependent",]

ggplot(datPlot2, aes(x=cntry,y=Freq,fill=factor(year))) +
        geom_bar(stat="identity",position="dodge") +
        theme(legend.position="top") +
        theme(axis.text.x = element_text(angle = 90))

```

### Education level (share of low education hh's)

```{rcompareIvEdu}
dat10 <- data.frame(prop.table(svytable(~edu2 + cntry, d.df10), 2) * 100)
names(dat10) <- c("Var","cntry","Freq")
dat10$year <- 2010
dat06 <- data.frame(prop.table(svytable(~edu2 + cntry, d.df06), 2) * 100)
names(dat06) <- c("Var","cntry","Freq")
dat06$year <- 2006
datPlot <- rbind(dat06,dat10)

datPlot2 <- datPlot[datPlot$Var == "no or compulsory",]

ggplot(datPlot2, aes(x=cntry,y=Freq,fill=factor(year))) +
        geom_bar(stat="identity",position="dodge") +
        theme(legend.position="top") +
        theme(axis.text.x = element_text(angle = 90))

```

### Income level (share of low income hh's)

```{rcompareIvInc}
dat10 <- data.frame(prop.table(svytable(~income2 + cntry, d.df10), 2) * 100)
names(dat10) <- c("Var","cntry","Freq")
dat10$year <- 2010
dat06 <- data.frame(prop.table(svytable(~income2 + cntry, d.df06), 2) * 100)
names(dat06) <- c("Var","cntry","Freq")
dat06$year <- 2006
datPlot <- rbind(dat06,dat10)

datPlot2 <- datPlot[datPlot$Var == "Low",]

ggplot(datPlot2, aes(x=cntry,y=Freq,fill=factor(year))) +
        geom_bar(stat="identity",position="dodge") +
        theme(legend.position="top") +
        theme(axis.text.x = element_text(angle = 90))

```

## Macro level

```{rcompareIvMacro, fig.height=16}
dfPlot <- macro[,c(1:2,13,14,17)]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("cntry","year"),
                value.vars=c("perGini",
                             "wbgi_vae",
                             "gdpChange"))
library(ggplot2)
ggplot(dfPlotW, aes(x=cntry,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))

```

# Associations

*Name is in the 2010 end of the line*

## Good governanance

### vs. Social blame

```{rcompareAssoGoodSB}
cbPalette <- c("#E69F00","#56B4E9","#999999","#009E73", "#F0E442", "#0072B2","#D55E00", "#CC79A7")
dfPlot <- macro[,c(1:4,13,14,16,17)]
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=wbgi_vae,y=socialBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

### vs. Individual blame

```{rcompareAssoGoodIB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=wbgi_vae,y=individualBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

## Change in GDP

### vs. Social blame

```{rcompareAssoGdpSB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=gdpChange,y=socialBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

### vs. Individual blame

```{rcompareAssoGdbIB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=gdpChange,y=individualBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

## Perceived Gini

### vs. Social blame

```{rcompareAssoPerginiSB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=perGini,y=socialBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=perGini,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=socialBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=perGini,y=socialBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=socialBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

### vs. Individual blame

```{rcompareAssoPerginiIB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=perGini,y=individualBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=perGini,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=individualBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=perGini,y=individualBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=individualBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```