### 2013-12-13 code/multilevel_models_test.Rmd
First pages commit

% Multilevel modelling

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/code/")
library(knitr)
knit("multilevel_models_test.Rmd")
system("~/.cabal/bin/pandoc multilevel_models_test.md --number-sections --template=../tex_templates/dcolumn.latex -o multilevel_models_test.tex")
#system("perl -pi -e 's/footnotesize/tiny/g' multilevel_models_test.tex")
system("pdflatex multilevel_models_test.tex")
```

```{rmultilevel_analysis, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, results='hide', eval=TRUE}
library(lme4.0)
#########################################
load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")

library(lme4)
summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
                      family = binomial, data=df))
summary(soc.1 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+(1|cntry),
                      family = binomial, data=df))
summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+(1|cntry),family = binomial, data=df))
summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
```

```{rstargazer, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(stargazer)
stargazer(soc.1,soc.2,soc.3,soc.4,
          title="testailua", ord.intercepts = TRUE)
```

```{rmemisc, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(memisc)
toLatex(mtable(soc.1,soc.2,soc.3,soc.4))
```

```{rtexreg, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(texreg)
texreg(list(soc.1,soc.2,soc.3,soc.4), use.packages=FALSE, bold=0.5)
#extract.mer(soc.1, include.pvalues=TRUE)
#summary(soc.1)
```