### 2014-03-28 index.Rmd
added tables to presentation

---
title: "Attributions for Poverty in Post-Socialist Countries"
author: [Markus Kainu & Mikko Niemelä]
---

****

# Introduction

This site gathers together all the data analysis behind the actual paper. Data analysis is broken down in three parts and each part is presented in .html format for human eyes and in text source including the R-code required for reproducing the analysis. As both the [micro data](http://www.ebrd.com/pages/research/publications/special/transitionII.shtml) and [macro data](http://www.qog.pol.gu.se/data/) are available as open data, anyone can rerun these analysis.

# Data Analysis

## Cleaning the data & variable summaries

| html-ouput | source code |
| ------------ | -------------- |
| [loading, cleaning and summarizing](loadClean.html) | [source](loadClean.Rmd) |

## Descriptive analysis

| html-ouput | source code |
| ------------ | -------------- |
| [Descriptive analysis](descriptive_analysis.html) | [source](descriptive_analysis.Rmd) |
| [Spatial visualization](mapPlots.html) | [source](mapPlots.Rmd) |
| [Graphics in the paper](finalPlots.html) | [source](finalPlots.Rmd) |

### Misc analysis

| html-ouput | source code |
| ------------ | -------------- |
| [Comparing waves 2006 and 2011](compareWaves.html) | [source](compareWaves.Rmd) |
| [Exploring the macro level correlations](correlations.html) | [source](correlations.md) |

## Modelling the data

| html-ouput | source code |
| ------------ | -------------- |
| [Regression analysis](regressionModelling.html) | [source](regressionModelling.Rmd) |

# Article draft

[Article draft in .pdf](article2013.pdf)

# Presentation slides

| html-ouput | source code |
| ------------ | -------------- |
| [slides](attrib_slides.html) | [source](attrib_slides.Rpres) |

---------

**Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.**