### 2014-03-29 mapPlots.Rmd
Ranska puuttui kuvailevistagit add . Lisäsin ja viilasin kuviot ja analyysit..

---
title: Spatial visualization
output:
  html_document:
    toc: true
    theme: flatly
---

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=FALSE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=5)

```

# Country grouping

```{rmapPlotsLoadData}

library(RColorBrewer)
library(mapproj)
library(maptools)
library(rgdal)
library(rgeos)
library(grid)
load("~/workspace/lits/clone/attributions/data/lits.RData")
map <- suppressWarnings(readOGR(dsn="/home/aurelius/data/shapefiles/lits/", 
                                layer="lits_all_simple_2014_crimea"))

map$NAME <- as.character(map$NAME)
map$NAME[map$NAME == "United Kingdom"] <- 'Great Britain'
map$group_analysis[map$NAME %in% c('France','Germany',
                                        'Great Britain',
                                        'Italy','Sweden')] <- 'Western Europe'

map$group_analysis[map$NAME %in% c('Czech Republic','Estonia',
                                        'Hungary','Bulgaria',
                                        'Latvia','Lithuania',
                                        'Poland','Slovakia',
                                        'Slovenia','Romania')] <- 'CEE'

map$group_analysis[map$NAME %in% c('Armenia','Azerbaijan',
                                        'Belarus','Georgia',
                                        'Kazakhstan','Kyrgyzstan',
                                        'Moldova','Tajikistan',
                                        'Ukraine','Uzbekistan','Russia')] <- 'CIS'
map$NAME <- as.factor(map$NAME)

# shapefile into data.frame
map@data$id <- rownames(map@data)
library(ggplot2)
map.points <- fortify(map, region="id")
map.df <- merge(map.points, map@data, by="id")
# only east of longitude -25
map.df <- subset(map.df, long >= -15 & long <= 80)
# names for countries in analysis (excluding others)
namedata <- map.df[!is.na(map.df$group_analysis),]
cnames <- stats:::aggregate.formula(cbind(long, lat) ~ NAME, data=namedata, mean)
cbPalette <- c("#999999", "#E69F00",
               "#56B4E9", "#009E73",
               "#F0E442", "#0072B2",
               "gray90", "#CC79A7")
```

```{rmapPlotsGroup1, fig.height=13, fig.width=16}
library(grid)
ggplot(map.df, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
  scale_fill_manual(values=cbPalette) +
  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
  coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.position="right")

```

# Dependent variables

**The darker the higher the support**

## Support for Social blame

```{rmapPlotsDV1}
### Faceted showing population shares of explanations
library(scales)
plotMap <- merge(map.df,macro,by.x="NAME",by.y="cntry")
plotMap <- plotMap[plotMap$year == 2010,]
plotMap <- plotMap[order(plotMap$order),]

ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = socialBlame)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(37,65)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +
  theme(legend.position="right") +
  theme(legend.text = element_text(size=16)) +
  scale_fill_gradient(low = "white", high = "orange")

```

## Individual blame

```{rmapPlotsDV2}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = individualBlame)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +
  theme(legend.position="right") +
  theme(legend.text = element_text(size=16)) +
  scale_fill_gradient(low = "white", high = "orange")
```

## Social fate

```{rmapPlotsDV3}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = socialFate)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +

  theme(legend.position="right") +
  theme(legend.text = element_text(size=16)) +
  scale_fill_gradient(low = "white", high = "orange")
```

## Individual fate

```{rmapPlotsDV4}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = individualFate)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +

  theme(legend.position="right") +
  theme(legend.text = element_text(size=16)) +
  scale_fill_gradient(low = "white", high = "orange")
```

## Not stated

```{rmapPlotsDV5}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = notStated)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +

  theme(legend.position="right") +
  theme(legend.text = element_text(size=16)) +
  scale_fill_gradient(low = "white", high = "orange")
```

## Dont know

```{rmapPlotsDV6}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = dontKnow)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +

  theme(legend.position="right") +
  theme(legend.text = element_text(size=16)) +
  scale_fill_gradient(low = "white", high = "orange")
```

# Independent variables

## Individual level

## Macro level

**Blue is "positive", red is "negative"**

### GDP change

```{rmapPlotsIvMacro1}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = gdpChange)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +
  scale_fill_gradient2(low=muted("red"), high=muted("green")) +
  theme(legend.position="right") +
  theme(legend.text = element_text(size=16))
```

### Human Development Index

```{rmapPlotsIvMacro2}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = undp_hdi)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +
  scale_fill_gradient2(low=muted("red"), high=muted("green"), midpoint=0.7570) +
  theme(legend.position="right") +
  theme(legend.text = element_text(size=16))
```

### Voice and Accountability

```{rmapPlotsIvMacro3}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_polygon(aes(fill = wbgi_vae)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
      coord_map(project="orthographic", xlim=c(-5,80),
                ylim=c(40,62)) +
  theme(legend.key.size = unit(10, "mm")) +
  theme(legend.title = element_text(size=16)) +
  scale_fill_gradient2(low=muted("red"), high=muted("green")) +
  theme(legend.position="right") +
  theme(legend.text = element_text(size=16))
```