### 2015-01-14 descriptive_analysis.Rmd
some polisihing for the analysis

---
title: "Descriptive visualizations"
output:
  html_document:
    number_sections: yes
    theme: flatly
    toc: yes
---

# Visualizing the predictors

## Perceived income

### Original variable

```{rdescplot1_1}
load("~/workspace/lits/clone/attributions/data/lits.RData")
library(ggplot2)
library(reshape2)

# 
# 
# df$cntry <- factor(df$cntry)
# tbl <- data.frame(prop.table(table(df$cntry,df$q227),1)*100)
# reg <- as.data.frame(table(df$cntry,df$group_general))
# reg <- subset(reg, Freq > 0)
# reg <- reg[with(reg, order(Var2)), ]
# x <- merge(tbl,reg[,1:2],by="Var1")
# names(x) <- c("cntry","cond","freq","group")
# 
# x <- x[with(x, order(cond)), ]
# x$cond <- as.numeric(levels(x$cond))[x$cond]
# x <- subset(x, cond >= -1)
# x$cntry <- factor(x$cntry, levels=reg$Var1)
# 
# ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
#   geom_bar(stat="identity") +
#   facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

#### Transformed variable

```{rdescplot1_2}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$income),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("Low","High"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Transfer dependency/Income source

### Original variable

```{rdescplot2_1, eval=FALSE}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q226m),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Transformed variable

```{rdescplot2_2}

df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$incsour3),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("dependent","notdependent"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Past

### Original variable

```{rdescplot3_1, eval=FALSE}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q228),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cond <- as.numeric(levels(x$cond))[x$cond]
x <- subset(x, cond >= -1)
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Transformed variable

```{rdescplot3_2}

tbl <- data.frame(prop.table(table(df$cntry,df$past),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("same or better","worse"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Future

### Original variable

```{rdescplot4_1, eval=FALSE}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q229),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cond <- as.numeric(levels(x$cond))[x$cond]
x <- subset(x, cond >= -1)
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Transformed variable

```{rdescplot4_2}

tbl <- data.frame(prop.table(table(df$cntry,df$future),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("same or better","worse"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Education

### Original variable

```{rdescplot5_1, eval=FALSE}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q515),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cond <- as.numeric(levels(x$cond))[x$cond]
x <- subset(x, cond >= -1)
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Transformed variable

```{rdescplot5_2}

tbl <- data.frame(prop.table(table(df$cntry,df$edu2),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("no or compulsory","higher"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Crises

### Original variable

```{rdescplot6_1, eval=FALSE}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q801),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  scale_fill_manual(values=c("red","orange","yellow","green","dim grey","grey"))

```

### Transformed variable

```{rdescplot6_2}

tbl <- data.frame(prop.table(table(df$cntry,df$crise),1)*100)
reg <- as.data.frame(table(df$cntry,df$group_general))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("Little or not at all","Great or fair amount"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

## Macro level

```{rdescplot7, eval=FALSE}
df <- macro[with(macro, order(perGini)), ]
#df$cntry <- factor(df$cntry, levels=df$cntry)
ggplot(df, aes(x=cntry,y=perGini, colour=group_general)) +
  geom_point(size=5) +
  labs(title="gini coeficient") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{rdescplot8}
df <- macro[with(macro, order(gdpChange)), ]
#df$cntry <- factor(df$cntry, levels=df$cntry)
ggplot(df, aes(x=cntry,y=gdpChange, colour=group_general)) +
  geom_point(size=5) +
  labs(title="gdp change 2007 - 2010") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{rdescplot9}
df <- macro[with(macro, order(wbgi_vae)), ]
#df$cntry <- factor(df$cntry, levels=df$cntry)

ggplot(df, aes(x=cntry,y=wbgi_vae, colour=group_general)) +
  geom_point(size=5) +
    labs(title="Voice and accountability") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{rdescplot10}
library(GGally)
ggpairs(macro[, c("perGini", "gdpChange", "wbgi_vae","socialBlame","individualBlame")])

```

# Dependent variable